Approaching Optimal Duplicate Detection in a Sliding Window

### 2020-06-22 article_content.tex
COCOON preparing camera-ready

In practice, additional constraints exist that we can capture with the following definition:
\begin{definition}[wDDP with bounded memory]
	At every time step $n$, given $e^{\star}$ and a current state (dependent on history) of at most $M$ bits, solve the wDDP for $E_n$ and $e^{\star}$.
\end{definition}
%Even when the stream is bounded, i.e. $\lim_{n\to\infty}|E_n| < \infty$, it may be impractical to perform a complete lookup and approximate solutions are desirable. 
%Perfect detection of duplicates in a stream with $U$ possible values requires $U$ bits of memory \cite[Theorem 2.1]{10.1145/3297280.3297335}: even for modest situations (e.g. 64-bit nonces, corresponding to $2^{64}$ possible values) this is quickly impractical. As a result, we need a further relaxation of the DDP that allows for errors: 
Perfect detection is however not always reachable and it might be more practical to work on a further relaxation of the problem, allowing for errors.

@@

Unfortunately, for $w > w_\text{max}$ the performance of SHF degrades very rapidly. We therefore turn our attention to existing data structures designed for the \enquote{non-windowed} setting. We show that some of them outperform dedicated data structures, including SHF, in the $w > w_\text{max}$ regime.

We then introduce the \enquote{queuing construction}, a black box transformation of non-windowed data structures into windowed ones, that improves their performance in the wDDP setting.

Finally, we provide an analysis of our queueing construction's resistance to adversarial streams.

\subsection{Related work}\label{sub:related}
The notion of sliding window was, as far as we know, first introduced in \cite{10.1145/1060745.1060753}.%; but several variations exist that are incomparable to one another (e.g. \cite{shtul2020agepartitioned}).
The wDDP formulation we rely on is due to \cite{Yoo10,She08}, which also introduce algorithms for solving the wDDP approximately.

The notion of using subfilters, as in the queuing construction, can be found in the A2 filter's design \cite{Yoo10} and a variation thereof can be found in \cite{shtul2020agepartitioned} but in a different DDP formulation. The A2 is built from two Bloom filters, a construction which we generalize and analyse generically in this paper. Similarly, the construction in \cite{shtul2020agepartitioned} only works with Bloom Filters.
Literature review collects the following DDFs \cite{Yoo10,Den06,10.1145/3297280.3297335,Dut13,She08,Fan14}, which we will consider in this paper . The structure proposed in \cite{10.1145/1060745.1060753} is not designed for wDDP but a variant called `landmark` sliding window, which consists of a zero-resetting of the memory at some user-defined epochs.
%\section{Duplicate Detection and Filter Saturation}
\label{sec:dudefisa}
\section{Notations and basic definitions}
We consider an unbounded stream $E = (e_1, e_2, \dotsc, e_n, \dotsc)$ with elements belonging to an alphabet $\Gamma$.

@@

For a filter, the false positive probability ($\FP^w_i$) is the probability that after $i$ insertions, the \emph{unseen} element $e_i\notwdup E$ is a false positive over $w$. The false positive rate $\FPR^w_i$ is the number of false positives divided by the number of unseen elements in E\footnote{We observe that $\FPR^w_n = \frac{1}{n}\sum_{i=1}^n \FP^w_i$, and similarly for $\FNR^w_n$.}. We similarly define the false negative probability $\FN^w_i$, and the false negative rate $\FNR^w_i$.

\paragraph{Remark.} For benchmarking, we usually measure the error rate $ER = \FPR^w + \FNR^w$, as it allows a practical ranking of the solutions. An error rate of $0$ means a perfect filter, while a filter answering randomly has an error rate of $1$. A filter being always wrong has an error rate of $2$.

@@

\begin{proof}
	Due to page limit restrictions, all theorem proofs are in Appendix~\ref{app:proofs}.
\end{proof}

@@

\begin{restatable}{thm}{shf}\label{thm:SHF}
	Let $w \in \mathbb N$. Let $M \simeq 2w\log_2w$, then the wDDP can be solved with almost no error using $M$ memory bits.

	More precisely, it is possible to create a filter of $M$ bits with an FN of $0$, an FP of $1 - (1-\frac1{w^2})^w \sim \frac 1w$, and a time complexity of $O(w)$.

@@

This gives us the Short Hash Filter (SHF), described in Algorithm~\ref{alg:SHF}. The implementation relies on a double-ended queue or a ring buffer, which allows pushing at beginning of a queue and popping at the end in constant time. A variant is described in Appendix~\ref{app:CSHF}.

\begin{algorithm}
	\caption{SHF \textsf{Setup}, \textsf{Lookup} and \textsf{Insert}}\label{alg:SHF}

	\begin{algorithmic}[1]
	\Function{Setup}{$M, w$}
	\Comment $M$ is the available memory, $w$ the size of the sliding window
	\State $h \gets$ hash function of codomain size $\lfloor \frac M{2w} - \frac 12 \log_2 w \rfloor$
	\State $Q \gets \emptyset$ \Comment $Q$ is a queue of elements of size $h$
	\State $D \gets \emptyset$ \Comment $D$ is a dictionary $h \Rightarrow$  counter (of max value $w$)
	\EndFunction 
	\end{algorithmic}

	\begin{multicols}{2}
	\begin{algorithmic}[1]
		\Function{Insert}{$e$}
		\State $Q.\mathsf{Push\_Front}(h(e))$
		\State $D[h(e)]$++

		\If{$Q.\mathsf{length}() > w$}
		\State $h' \gets Q.\mathsf{Pop\_back}()$
		\State $D[h']$-{}-
		\If{$D[h'] = 0$}
		\State Erase key $D[h']$
		\EndIf
		\EndIf
		\EndFunction
	\end{algorithmic}

	\begin{algorithmic}[1]
	\item[]
	\Function{Lookup}{$e$}
	\If{$D[h(e)] > 0$}
	\State \Return \DUPLICATE
	\Else
	\State \Return \UNSEEN
	\EndIf
	\EndFunction
	\item[]
\end{algorithmic}

	\end{multicols}
\end{algorithm}

@@

%\paragraph{Remark:} A CSHF of size $M$ has the same error rate than an SHF of size $2M + w\log_2 w$.
\paragraph{Saturation.}
SHF has strictly increasing error probabilities, which reach a threshold of $1/2$ for some maximum window size $w_\text{max}$. Beyond this value, these filters saturate extremely quickly: in other words, most SHF will either have an error rate of $0$ or $1$.

An illustration of this phenomenon can be seen in Figure~\ref{fig:shorts}, 
which shows the error rates for SHF with $M=10^5$, against 
a uniformly random stream of $18$-bit elements ($|\Gamma| = 2^{18}$). The benchmark used a finite stream of length $10^6$.

@@

In particular, the asymptotic error probability $EP_\infty$ satisfies

$
	EP_\infty \geq \left(1 - \frac{1}{|\Gamma|}\right)^{M} \approx 1 - M/|\Gamma|.
$

@@

\subsection{Saturation Resistance of DDFs}\label{sub:saturation}
We now evaluate the saturation rate for several DDFs, in the original DDP setting (without sliding window). Parameters are chosen to yield equivalent memory footprints and were taken from \cite{10.1145/3297280.3297335}, namely:
%
\begin{compactitem}
	\item QHT \cite{10.1145/3297280.3297335}, 1 bucket per row, 3 bits per fingerprint;
	\item SQF \cite{Dut13}, 1 bucket per row, $r = 2$ and $r' = 1$;
	\item Cuckoo Filter \cite{Fan14}, cells containing 1 element of 3 bits each;
	\item Stable Bloom Filter (SBF) \cite{Den06}, 2 bits per cell, 2 hash functions, targeted FPR of 0.02.
%	\item $A2$ Filter \cite{Yoo10}, targeted FPR of $0.1$ on the sliding window.
%	\item Block-Decaying Bloom Filter (b\_DBF) \cite{She08}, sliding window of  elements.
\end{compactitem}
%

@@

\subsection{Performance in wDDP}
We now consider the performance of the filters just discussed in the \emph{windowed} setting, for which they were \emph{not} designed. In particular, it is not possible to adjust their parameters as a function of $w$.

Remarkably, some of these filters still outperform dedicated windowed filters for some window sizes at least, as shown in Figure~\ref{fig:nwddf-in-wddp}. In this benchmark, we used the following filters:
\begin{itemize}
	\item block decaying Bloom Filter\footnote{Note that by design, a b\_DBF of $10^5$ bits cannot operate for $w > 6000$.} (b\_DBF) \cite{She08}, sliding window of size $w$
	\item A2 filter \cite{Yoo10}, changing subfilter every $w/2$ insertions
	\item QHT \cite{10.1145/3297280.3297335}, 1 bucket per row, $3$ bits per fingerprint
\end{itemize}

@@

\paragraph{Principle of operation.}
Let $\mathcal F$ be a DDF. Rather than allocating the whole memory to $\mathcal F$, we will create $L$ copies of $\mathcal F$, each using a fraction of the available memory. Each of these \emph{subfilters} has a limited timespan, and is allowed up to $c$ insertions. The subfilters are organised in a queue.

When inserting a new element in the queuing filter, it is inserted in the topmost subfilter of the queue. After $c$ insertions, a new empty filter is added to the queue, and the oldest subfilter is popped and erased.

@@

\paragraph{Pseudocode.}
We give a brief pseudocode for the queuing filter's functions \textsf{Lookup} and \textsf{Insert}, as well as a \textsf{Setup} function for initialisation, in Algorithm~\ref{alg:queuing}. We introduced for simplicity a constructor $\mathcal F.\textsf{Setup}$ that takes as input an integer $M$ and outputs an initialized empty filter $\mathcal F$ of size at most $M$. Here \texttt{subfilters} is a FIFO that has a \texttt{pop} and \texttt{push\_first} operations, which respectevely removes the last element in the queue or inserts a new item in first position.

\begin{algorithm}[]
	\caption{Queuing Filter \textsf{Setup}, \textsf{Lookup} and \textsf{Insert}}\label{alg:queuing}

	\begin{algorithmic}[1]
	\Function{Setup}{$\mathcal F, M, L, c$}
		\Comment $M$ is the available memory, $\mathcal F$ the subfilter structure, $L$ the number of subfilters and $c$ the number of insertions per subfilter
		\State subfilters $\gets \emptyset$
		\State counter $\gets 0$
		\State $m \gets \lfloor M/L \rfloor$
		\For{$i$ from $0$ to $L-1$}
			\State subfilters.push\_first$(\mathcal F.\mathsf{Setup}(m))$
		\EndFor
		\State \textbf{store} (subfilters, $L$, $m$, counter)
%		\State\Return filter
	\EndFunction 
	\end{algorithmic}

	\begin{multicols}{2}
	\begin{algorithmic}[1]
	\Function{Lookup}{$e$}
		\For{$i$ from $0$ to $L-1$}
			\If{subfilters[i]$.\mathsf{Lookup(e)}$}
				\State \Return $\DUPLICATE$			
			\EndIf
		\EndFor
		\State \Return \UNSEEN
	\EndFunction
	\item[]
	\end{algorithmic}

	\begin{algorithmic}[1]
	\Function{Insert}{$e$}
		\State subfilters[0]$.\mathsf{Insert}(e)$
		\State counter$++$
		\If{counter == $c$}
			\State subfilters.pop()
			\State subfilters.push\_first($\mathcal F.\textsf{Setup}(m)$)
		\EndIf
	\EndFunction
	\end{algorithmic}
	\end{multicols}
\end{algorithm}

@@

Hence, if $w < cL$, the queuing filter can hold information about \emph{more than $w$ elements}.

@@

\begin{restatable}{thm}{pfpQueue}
	\label{thm:queue-pfp}
	 Let $\FP_{\mathcal Q, m}^w$ be the false positive probability%(def. \ref{def:pfp-pfn})
	 of $\mathcal Q$ after $m > w$ insertions, over a sliding window of size $w = cL$, we have \\
	$$%\begin{equation}
		\label{eq:queue-pfp}
		\FP_{\mathcal Q, m}^w = 1 - \left(1 - \FP_{\mathcal F,c}\right)^{L-1} \left(1 - \FP_{\mathcal F, m \bmod c }\right)
	$$ %\end{equation}
	where %$\bmod$ is the modulo operator and 
	 $\FP_{\mathcal F,m}$ is the false positive probability of a subfilter $\mathcal F$ after $m$ insertions.
\end{restatable}

@@

\subsubsection{False Negative Probability}\label{sub:fnr}
\begin{restatable}{thm}{pfnQueue}
	\label{thm:queue-pfn}
	Let $\FN_{\mathcal Q, m}^w$ be the false negative probability of $\mathcal Q$ after $m > w$ insertions on a sliding window of size $w = cL$, we have
	$$%\begin{equation}
	\FN_{\mathcal Q, m}^w 
	=  u_c^{L-1} u_{m \bmod c}
	$$%\end{equation}
	where %$\bmod$ is the modulo operator, and
	we have introduced the short-hand notation $u_\eta= p_{\eta}\FN_{\mathcal F,\eta} + \left(1-p_{\eta}\right)\left(1-\FP_{\mathcal F,\eta}\right)$ where 
	$\FN_{\mathcal F,\eta}$ (resp. $\FP_{\mathcal F, \eta}$) is the false negative probability (resp. false positive) of the subfilter $\mathcal F$ after $\eta$ insertions, and
	$%\begin{equation*}
		p_\eta = \frac{1-\left(1-\frac 1{|\Gamma|}\right)^\eta}{1-\left(1-\frac 1{|\Gamma|}\right)^w} \approx \frac{\eta}{w}.
	$%\end{equation*}
\end{restatable}

@@

\subsection{FNR and FPR}
From the above expressions we can derive relatively compact explicit formulas for the queuing filter's FPR and FNR when $m = cn$ for $n$ a positive integer.
\begin{restatable}{thm}{fprQueue}
	Let $\FPR_{\mathcal Q, m}^w$ be the false positive rate of $\mathcal Q$ after $m=cn > w$ insertions on a sliding window of size $w = cL$, we have
	\begin{equation*}
	\FPR_{\mathcal Q, cn}^w = 1 -  \frac{(1 - \FP_{\mathcal F, c})^{L-1}}{c}\sum_{\ell = 0}^{c-1} (1-\FP_{\mathcal F, \ell}).
	\end{equation*}
\end{restatable}
%\begin{proof}
%
%\begin{align*}
%	\FPR_{\mathcal Q, cn}^w
%	& = \frac{1}{cn} \sum_{k=1}^{cn} \FP_{\mathcal Q, k}^w 
%	 = \frac{1}{cn} \sum_{k=1}^{n} \sum_{\ell = 0}^{c-1} \FP_{\mathcal Q, k + \ell}^w 
%	 = \frac1c \sum_{\ell = 0}^{c-1} \FP_{\mathcal Q, \ell}^w \\
%	& = \frac1c \sum_{\ell = 0}^{c-1}  1 - (1 - \FP_{\mathcal F, c})^{L-1}(1 - \FP_{\mathcal F}, \ell) \\
%	& = 1 - \frac1c(1 - \FP_{\mathcal F, c})^{L-1}\sum_{\ell = 0}^{c-1} (1-\FP_{\mathcal F, \ell})
%\end{align*}
%\end{proof}

\begin{restatable}{thm}{fnrQueue}
	Let $\FNR_{\mathcal Q, m}^w$ be the false negative rate of $\mathcal Q$ after $m=cn > w$ insertions on a sliding window of size $w = cL$, we have
	\begin{equation*}
	\FNR_{\mathcal Q, cn}^w = \frac{u_c^{L-1}}{c}  \sum_{\ell = 0}^{c-1}u_\ell.
	\end{equation*}
\end{restatable}

@@

Because $u_\eta = p_{\eta}\FN_{\mathcal F,\eta} + \left(1-p_{\eta}\right)\left(1-\FP_{\mathcal F,\eta}\right)$, differentiating with respect to $L$, knowing that $w = Lc$, and equating the derivative to $0$, one can find the optimal value for $L$ by solving for $x$, which has been obtained via Mathematica:
%\begin{equation*}

\begin{align*}
&-\alpha \beta^{-1+x} \log(\beta) 
+ \left(\beta + \FN_{\mathcal F, c} (-1+x)\right)^{-2+x} x^{-x} \left[\vphantom{\left(\FN^{\FN}\right)}\right. 
	-\alpha \beta + \FN_{\mathcal F, c} \left(-\beta (-2+x)+ \FN_{\mathcal F, c} (-1+x)\right) \\
&\quad	+ \left(\alpha+ \FN_{\mathcal F, c} (-1+x)\right)
\times\left(\beta + \FN_{\mathcal F, c} (-1+x)\right) \left(\log\left(\beta + \FN_{\mathcal F, c} (-1+x)\right)-\log(x)\right)
\left.\vphantom{\left(\FN^{\FN}\right)}\right] = 0
\end{align*}

@@

Indeed, a simple symmetry argument shows that a heterogenous selection of subfilters is always worse than a homogeneous one: the crux is that all subfilters play the same role in turn. Therefore we lose nothing by replacing atomically one subfilter by a more efficient one. Applying this to each subfilter we end up with a homogenous selection.

@@

This section provides details and additional information on the benchmarking experiments run to validate the above analysis. All code is accessible online and will be disclosed after peer review.

@@

\begin{restatable}[Bounds on false negative resistance]{thm}{fnrRes}
	Let $\mathcal Q$ be a filter of $L$ subfilters of kind $\mathcal F$, with $c$ insertions maximum per subfilter, and let $w$ be a sliding window.

	If $\mathcal F$ is $(p, c)$-resistant to adversarial false negative attacks, then $\mathcal A$ can win the adversarial game on the sliding window $w$ with probability at least $p^L$.

>>>>

In practice, additional constraints exist that we can capture with the following definition:
\begin{definition}[wDDP with bounded memory]
	At every time step $n$, given $e^{\star}$ and a current state (dependent on history) of at most $M$ bits, solve the wDDP for $E_n$ and $e^{\star}$.
\end{definition}

A filter attempting to solve the wDDP is called a duplicate detection filter (DDF).
%Even when the stream is bounded, i.e. $\lim_{n\to\infty}|E_n| < \infty$, it may be impractical to perform a complete lookup and approximate solutions are desirable. 
%Perfect detection of duplicates in a stream with $U$ possible values requires $U$ bits of memory \cite[Theorem 2.1]{10.1145/3297280.3297335}: even for modest situations (e.g. 64-bit nonces, corresponding to $2^{64}$ possible values) this is quickly impractical. As a result, we need a further relaxation of the DDP that allows for errors: 
Perfect detection is however not always reachable and it might be more practical to work on a further relaxation of the problem, allowing for errors.

@@

Unfortunately, for $w > w_\text{max}$ the performance of SHF degrades very rapidly. We therefore turn our attention to existing data structures designed for the \enquote{non-windowed} setting. We show that some of them outperform dedicated data structures, including SHF, in the $w > w_\text{max}$ regime. However, we exhibit a lower bound on the error rate of such filters.

We then introduce the \enquote{queuing construction}, a black box transformation of non-windowed data structures into windowed ones, that improves their performance in the wDDP setting.
Finally, we provide an analysis of our queueing construction's resistance to adversarial streams.

\subsection{Related work}\label{sub:related}
The notion of sliding window was, as far as we know, first introduced in \cite{10.1145/1060745.1060753}.%; but several variations exist that are incomparable to one another (e.g. \cite{shtul2020agepartitioned}).
The wDDP formulation we rely on is due to \cite{Yoo10,She08}, which also introduce algorithms for solving the wDDP approximately. To the best of our knowledge, the discrepancy between the wDDP and the DDP and the wDDP has not been studied. Similarly, the relation between $w$ (the sliding window size), the error rate and $M$ (the available memory) is not always present in similar papers, one exception being \cite{She08}.

The notion of using subfilters, as in the queuing construction, can be found in the A2 filter's design \cite{Yoo10} and a variation thereof can be found in \cite{shtul2020agepartitioned} but in a different DDP formulation. The A2 is built from two Bloom filters, a construction which we generalize and analyse generically in this paper. Similarly, the construction in \cite{shtul2020agepartitioned} only works with Bloom Filters.
Literature review collects the following DDFs \cite{Yoo10,Den06,10.1145/3297280.3297335,Dut13,She08,Fan14}, which we will consider in this paper . The structure proposed in \cite{10.1145/1060745.1060753} is not designed for wDDP but a variant called `landmark` sliding window, which consists of a zero-resetting of the memory at some user-defined epochs.
%\section{Duplicate Detection and Filter Saturation}
\label{sec:dudefisa}
\section{Notations and basic definitions}
We consider an unbounded stream $E = (e_1, e_2, \dotsc, e_n, \dotsc)$ of elements from an alphabet $\Gamma$.

@@

The filter false positive probability ($\FP^w_i$) is the probability that after $i$ insertions, the \emph{unseen} element $e_i\notwdup E$ is a false positive over $w$. The false positive rate $\FPR^w_i$ is the number of false positives divided by the number of unseen elements in E\footnote{We observe that $\FPR^w_n = \frac{1}{n}\sum_{i=1}^n \FP^w_i$, and similarly for $\FNR^w_n$.}. We similarly define the false negative probability $\FN^w_i$, and the false negative rate $\FNR^w_i$.

\paragraph{Remark.} For benchmarking, we usually measure the error rate $ER^w = \FPR^w + \FNR^w$, as it allows a practical ranking of the solutions. An error rate of $0$ means a perfect filter, while a filter answering randomly has an error rate of $1$. A filter being always wrong has an error rate of $2$.

@@

\begin{proof}
	Due to page limit restrictions, all proofs are in Appendix~\ref{app:proofs}.
\end{proof}

@@

\begin{restatable}{thm}{shf}\label{thm:SHF}
	Let $w \in \mathbb N$. Let $M \simeq 2w\log_2w$, then the wDDP can be solved with almost no error using $M$ memory bits.
	More precisely, it is possible to create a filter of $M$ bits with an FN of $0$, an FP of $1 - (1-\frac1{w^2})^w \sim \frac 1w$, and a time complexity of $O(w)$.

@@

This gives us the Short Hash Filter (SHF), described in Algorithm~\ref{alg:SHF}, in Appendix~\ref{app:algo}. The implementation relies on a double-ended queue or a ring buffer, which allows pushing at beginning of a queue and popping at the end in constant time. A variant is described in Appendix~\ref{app:CSHF}.

%\begin{algorithm}
%	\caption{SHF \textsf{Setup}, \textsf{Lookup} and \textsf{Insert}}\label{alg:SHF}
%	
%	\begin{algorithmic}[1]
%	\Function{Setup}{$M, w$}
%	\Comment $M$ is the available memory, $w$ the size of the sliding window
%	\State $h \gets$ hash function of codomain size $\lfloor \frac M{2w} - \frac 12 \log_2 w \rfloor$
%	\State $Q \gets \emptyset$ \Comment $Q$ is a queue of elements of size $h$
%	\State $D \gets \emptyset$ \Comment $D$ is a dictionary $h \Rightarrow$  counter (of max value $w$)
%	\EndFunction 
%	\end{algorithmic}
%
%	\begin{multicols}{2}
%	\begin{algorithmic}[1]
%		\Function{Insert}{$e$}
%		\State $Q.\mathsf{Push\_Front}(h(e))$
%		\State $D[h(e)]$++
%		
%		\If{$Q.\mathsf{length}() > w$}
%		\State $h' \gets Q.\mathsf{Pop\_back}()$
%		\State $D[h']$-{}-
%		\If{$D[h'] = 0$}
%		\State Erase key $D[h']$
%		\EndIf
%		\EndIf
%		\EndFunction
%	\end{algorithmic}	
%		
%	\begin{algorithmic}[1]
%	\item[]
%	\Function{Lookup}{$e$}
%	\If{$D[h(e)] > 0$}
%	\State \Return \DUPLICATE
%	\Else
%	\State \Return \UNSEEN
%	\EndIf
%	\EndFunction
%	\item[]
%\end{algorithmic}
%
%
%	\end{multicols}
%\end{algorithm}

@@

%\paragraph{Remark:} A CSHF of size $M$ has the same error rate than an SHF of size $2M + w\log_2 w$.
\paragraph{Saturation.}
SHF has strictly increasing error probabilities, which reach a threshold of $1/2$ for some maximum window size $w_\text{max}$. Beyond this value, these filters saturate extremely quickly: in other words, most SHF will either have an error rate of $0$ or $1$.
An illustration of this phenomenon can be seen in Figure~\ref{fig:shorts}, 
which shows the error rates for SHF with $M=10^5$, against 
a uniformly random stream of $18$-bit elements ($|\Gamma| = 2^{18}$). The benchmark used a finite stream of length $10^6$.

@@

The asymptotic error probability $EP_\infty$ satisfies 
$
	EP_\infty \geq \left(1 - \frac{1}{|\Gamma|}\right)^{M} \approx 1 - M/|\Gamma|.
$

@@

\subsection{Saturation Resistance of DDFs}\label{sub:saturation}
We now evaluate the saturation rate for several DDFs, in the original DDP setting (without sliding window). Parameters are chosen to yield equivalent memory footprints and were taken from \cite{10.1145/3297280.3297335}, namely:
%
\begin{compactitem}
	\item QHT \cite{10.1145/3297280.3297335}, 1 bucket per row, 3 bits per fingerprint;
	\item SQF \cite{Dut13}, 1 bucket per row, $r = 2$ and $r' = 1$;
	\item Cuckoo Filter \cite{Fan14}, cells containing 1 element of 3 bits each;
	\item Stable Bloom Filter (SBF) \cite{Den06}, 2 bits per cell, 2 hashes, targeted FPR of 0.02.
%	\item $A2$ Filter \cite{Yoo10}, targeted FPR of $0.1$ on the sliding window.
%	\item Block-Decaying Bloom Filter (b\_DBF) \cite{She08}, sliding window of  elements.
\end{compactitem}
%

@@

\subsection{Performance in wDDP}
We now consider the performance of the filters just discussed in the \emph{windowed} setting, for which they were \emph{not} designed. In particular, they cannot adjust their parameters as a function of $w$.
Remarkably, some of these filters still outperform dedicated windowed filters for some window sizes at least, as shown in Figure~\ref{fig:nwddf-in-wddp}. In this benchmark, we used the following filters:
\begin{compactitem}
	\item block decaying Bloom Filter\footnote{Note that by design, a b\_DBF of $10^5$ bits cannot operate for $w > 6000$.} (b\_DBF) \cite{She08}, sliding window of size $w$
	\item A2 filter \cite{Yoo10}, changing subfilter every $w/2$ insertions
	\item QHT \cite{10.1145/3297280.3297335}, 1 bucket per row, $3$ bits per fingerprint
\end{compactitem}

@@

\paragraph{Principle of operation.}
Let $\mathcal F$ be a DDF. Rather than allocating the whole memory to $\mathcal F$, we will create $L$ copies of $\mathcal F$, each using a fraction of the available memory. Each of these \emph{subfilters} has a limited timespan, and is allowed up to $c$ insertions. The subfilters are organised in a queue.
When inserting a new element in the queuing filter, it is inserted in the topmost subfilter of the queue. After $c$ insertions, a new empty filter is added to the queue, and the oldest subfilter is popped and erased.

@@

\paragraph{Pseudocode.}
We give a brief pseudocode for the queuing filter's functions \textsf{Lookup} and \textsf{Insert}, as well as a \textsf{Setup} function for initialisation, in Algorithm~\ref{alg:queuing}, in Appendix~\ref{app:algo}. We introduced for simplicity a constructor $\mathcal F.\textsf{Setup}$ that takes as input an integer $M$ and outputs an initialized empty filter $\mathcal F$ of size at most $M$. Here \texttt{subfilters} is a FIFO that has a \texttt{pop} and \texttt{push\_first} operations, which respectevely removes the last element in the queue or inserts a new item in first position.

%\begin{algorithm}[]
%	\caption{Queuing Filter \textsf{Setup}, \textsf{Lookup} and \textsf{Insert}}\label{alg:queuing}
%	\begin{algorithmic}[1]
%	\Function{Setup}{$\mathcal F, M, L, c$}
%		\Comment $M$ is the available memory, $\mathcal F$ the subfilter structure, $L$ the number of subfilters and $c$ the number of insertions per subfilter
%		\State subfilters $\gets \emptyset$
%		\State counter $\gets 0$
%		\State $m \gets \lfloor M/L \rfloor$
%		\For{$i$ from $0$ to $L-1$}
%			\State subfilters.push\_first$(\mathcal F.\mathsf{Setup}(m))$
%		\EndFor
%		\State \textbf{store} (subfilters, $L$, $m$, counter)
%%		\State\Return filter
%	\EndFunction 
%	\end{algorithmic}
%
%	\begin{multicols}{2}
%	\begin{algorithmic}[1]
%	\Function{Lookup}{$e$}
%		\For{$i$ from $0$ to $L-1$}
%			\If{subfilters[i]$.\mathsf{Lookup(e)}$}
%				\State \Return $\DUPLICATE$			
%			\EndIf
%		\EndFor
%		\State \Return \UNSEEN
%	\EndFunction
%	\item[]
%	\end{algorithmic}
%
%	\begin{algorithmic}[1]
%	\Function{Insert}{$e$}
%		\State subfilters[0]$.\mathsf{Insert}(e)$
%		\State counter$++$
%		\If{counter == $c$}
%			\State subfilters.pop()
%			\State subfilters.push\_first($\mathcal F.\textsf{Setup}(m)$)
%		\EndIf
%	\EndFunction
%	\end{algorithmic}
%	\end{multicols}
%\end{algorithm}

@@

If $w < cL$, the queuing filter can hold information about \emph{more than $w$ elements}.

@@

\begin{restatable}{thm}{pfpQueue}
	\label{thm:queue-pfp}
	 Let $\FP_{\mathcal Q, m}^w$ be the false positive probability%(def. \ref{def:pfp-pfn})
	 of $\mathcal Q$ after $m > w$ insertions, over a sliding window of size $w = cL$, we have
	$%\begin{equation}
		\label{eq:queue-pfp}
		\FP_{\mathcal Q, m}^w = 1 - \left(1 - \FP_{\mathcal F,c}\right)^{L-1} \left(1 - \FP_{\mathcal F, m \bmod c }\right)
	$ %\end{equation}
	where %$\bmod$ is the modulo operator and 
	 $\FP_{\mathcal F,m}$ is the false positive probability of a subfilter $\mathcal F$ after $m$ insertions.
\end{restatable}

@@

\subsubsection{False Negative Probability}\label{sub:fnr}
\begin{restatable}{thm}{pfnQueue}
	\label{thm:queue-pfn}
	Let $\FN_{\mathcal Q, m}^w$ be the false negative probability of $\mathcal Q$ after $m > w$ insertions on a sliding window of size $w = cL$, we have
	$%\begin{equation}
	\FN_{\mathcal Q, m}^w 
	=  u_c^{L-1} u_{m \bmod c}
	$%\end{equation}
	where %$\bmod$ is the modulo operator, and
	we have introduced the short-hand notation $u_\eta= p_{\eta}\FN_{\mathcal F,\eta} + \left(1-p_{\eta}\right)\left(1-\FP_{\mathcal F,\eta}\right)$ where 
	$\FN_{\mathcal F,\eta}$ (resp. $\FP_{\mathcal F, \eta}$) is the false negative probability (resp. false positive) of the subfilter $\mathcal F$ after $\eta$ insertions, and
	$%\begin{equation*}
		p_\eta = \frac{1-\left(1-\frac 1{|\Gamma|}\right)^\eta}{1-\left(1-\frac 1{|\Gamma|}\right)^w} \approx \frac{\eta}{w}.
	$%\end{equation*}
\end{restatable}

@@

\subsection{FNR and FPR}
From the above expressions we can derive relatively compact explicit formulas for the queuing filter's FPR and FNR when $m = cn$ for $n$ a positive integer.
\begin{restatable}{thm}{fprQueue}
	Let $\FPR_{\mathcal Q, m}^w$ be the false positive rate of $\mathcal Q$ after $m=cn > w$ insertions on a sliding window of size $w = cL$, we have
	$%\begin{equation*}
	\FPR_{\mathcal Q, cn}^w = 1 -  \frac{(1 - \FP_{\mathcal F, c})^{L-1}}{c}\sum_{\ell = 0}^{c-1} (1-\FP_{\mathcal F, \ell}).
	$%\end{equation*}
\end{restatable}
%\begin{proof}
%
%\begin{align*}
%	\FPR_{\mathcal Q, cn}^w
%	& = \frac{1}{cn} \sum_{k=1}^{cn} \FP_{\mathcal Q, k}^w 
%	 = \frac{1}{cn} \sum_{k=1}^{n} \sum_{\ell = 0}^{c-1} \FP_{\mathcal Q, k + \ell}^w 
%	 = \frac1c \sum_{\ell = 0}^{c-1} \FP_{\mathcal Q, \ell}^w \\
%	& = \frac1c \sum_{\ell = 0}^{c-1}  1 - (1 - \FP_{\mathcal F, c})^{L-1}(1 - \FP_{\mathcal F}, \ell) \\
%	& = 1 - \frac1c(1 - \FP_{\mathcal F, c})^{L-1}\sum_{\ell = 0}^{c-1} (1-\FP_{\mathcal F, \ell})
%\end{align*}
%\end{proof}

\begin{restatable}{thm}{fnrQueue}
	Let $\FNR_{\mathcal Q, m}^w$ be the false negative rate of $\mathcal Q$ after $m=cn > w$ insertions on a sliding window of size $w = cL$, we have
	$%\begin{equation*}
	\FNR_{\mathcal Q, cn}^w = \frac{u_c^{L-1}}{c}  \sum_{\ell = 0}^{c-1}u_\ell.
	$%\end{equation*}
\end{restatable}

@@

Because $u_\eta = p_{\eta}\FN_{\mathcal F,\eta} + \left(1-p_{\eta}\right)\left(1-\FP_{\mathcal F,\eta}\right)$, differentiating with respect to $L$, knowing that $w = Lc$, and equating the derivative to $0$, one can find the optimal value for $L$ by solving for $x$, which has been obtained via Mathematica:
%\begin{equation*}
\begin{align*}
&-\alpha \beta^{-1+x} \log(\beta) 
+ \left(\beta + \FN_{\mathcal F, c} (-1+x)\right)^{-2+x} x^{-x} \left[\vphantom{\left(\FN^{\FN}\right)}\right. 
	-\alpha \beta + \FN_{\mathcal F, c} \left(-\beta (-2+x)+ \FN_{\mathcal F, c} (-1+x)\right) \\
&\quad	+ \left(\alpha+ \FN_{\mathcal F, c} (-1+x)\right)
\times\left(\beta + \FN_{\mathcal F, c} (-1+x)\right) \left(\log\left(\beta + \FN_{\mathcal F, c} (-1+x)\right)-\log(x)\right)
\left.\vphantom{\left(\FN^{\FN}\right)}\right] = 0
\end{align*}

@@

Indeed, a simple symmetry argument shows that a heterogenous selection of subfilters is always worse than a homogeneous one: the crux is that all subfilters play the same role in turn. Therefore we lose nothing by replacing atomically one subfilter by a more efficient one. Applying this to each subfilter we end with a homogenous selection.

@@

This section provides details and additional information on the benchmarking experiments run to validate the above analysis. Our code is accessible online\footnote{\url{https://github.com/mariuslp/duplicate_sliding_benchmark}}.

@@

\begin{restatable}[Bounds on false negative resistance]{thm}{fnrRes}
	Let $\mathcal Q$ be a filter of $L$ subfilters of kind $\mathcal F$, with $c$ insertions maximum per subfilter, and let $w$ be a sliding window. 
	If $\mathcal F$ is $(p, c)$-resistant to adversarial false negative attacks, then $\mathcal A$ can win the adversarial game on the sliding window $w$ with probability at least $p^L$.
